CHAPTER TWENTY UNIFORM APROXIMATION BY POLYNOMIALS Introduction 20.1 Let $K^{0}$ be the interior of a compact set $\textstyle K$ in the complex plane.(By defini tion, $K^{0}$ is the union of all open discs which are subsets of $K{\mathrm{:}}$ ； of course, $K^{0}$ may be empty even if $ .K$ is not.) Let $P(K)$ denote the set of all functions on ${\cal K}\,\,\,\,\,\,\,\,\,\,{\cal K}$ which are uniform limits of polynomials in z. Which functions belong to $P(K){\dot{Y}}$ Two necessary conditions come to mind immediately：If fe P(K),then $f\in C(K)$ and $f\in H(K^{0}).$ The question arises whether these necessary conditions are also sufficient The answer is negative whenever ${\boldsymbol{K}}$ separates the plane （i.e., when the com plement of ${\cal K}\,\,\,\,$ is not connected). We saw this in Sec. 13.8. On the other hand, if ${\cal K}\,\,\,\,$ is an interval on the real axis Gin which case $K^{0}={\emptyset}),$ the Weierstrass approx imation theorem asserts that $$ P(K)=C(K). $$ So the answer is positive if $\displaystyle K$ is an interval. Runge's theorem also points in this direction, since it states, for compact sets K $\textstyle K$ which do not separate the plane, that $P(K)$ contains at least all those f∈ C(K) which have holomorphic extensions to some open set $\scriptstyle\mathbf{a}\gg\kappa$ In this chapter we shall prove the theorem of Mergelyan which states without any superfluous hypotheses, that the above-mentioned necessary condi- tions are also sufficient if ${\cal K}\,\,\,\,$ does not separate the plane The principal ingredients of the proof are： Tietze's extension theorem,a smoothing process invoving convolutions,Runge's theorem, and Lemma 20.2. whose proof depends on properties of the class ${\mathcal{P}}$ which was introduced in Chap 14. 386UNIFORM APPROXIMATION BY POLYNOMIALS 387 Some Lemmas 20.2 Lemma Suppose $D\!\!\!\!/$ is an open disc of radius r > 0, $\scriptstyle{E\cdot D}$ ${\boldsymbol{E}}$ is compact and connected, $\scriptstyle\mathbf{a}={\mathcal{S}}^{2}-E$ is connected, and the diameter of $\boldsymbol{E}$ is at least ${\boldsymbol{r}}.$ Then there is a function $g\in H(\Omega)$ and a constant b, with the following property: ${\cal I}{\cal f}$ $$ Q(\zeta,z)=g(z)+(\zeta-b)g^{2}(z), $$ (1） the inequalities $$ \begin{array}{c}{{|Q(\zeta,z)|<{\frac{100}{r}}}}\\ {{\Big|Q(\zeta,z)-{\frac{1}{z-\zeta}}\Big|<{\frac{1,000r^{2}}{1z-\zeta!}}}}\end{array} $$ (2) (3) hold for all $\varepsilon\in\Omega$ and for $a l l\,\zeta\in D,$ We recall that ${\boldsymbol{S}}^{2}$ is the Riemann sphere and that the diameter of $\boldsymbol{E}$ is the supremum of the numbers $|z_{1}-z_{2}|.$ where $z_{1}\in E{\mathrm{~and~}}z_{2}\in E.$ PROOF We assume, without loss of generality, that the center of $D\!\!\!\!/$ is at the origin. So $D=D(0;r)$ The implication $(d)\to(b)$ of Theorem 13.11 shows that such that $F(0)=\infty$ ${\mathbf{}}F$ has an fore a conformal mapping ${\mathbf{}}F$ of $U$ onto $\Omega$ $\Omega$ is simply con- nected. (Note that $\infty\in\Omega\}$ By the Riemann mapping theorem there is there- expansion of the form $$ F(w)={\frac{a}{w}}+\sum_{n=0}^{\infty}c_{n}w^{n}\qquad(w\in U). $$ (4） We define $$ g(z)=\frac{1}{a}\,F^{-1}(z)\qquad(z\in\Omega), $$ (5) where ${\boldsymbol{F}}^{\prime\prime}$ is the mapping of $\Omega$ onto $U$ which inverts ${\boldsymbol{F}},$ and we put $$ b={\frac{1}{2\pi i}}\ {\biggr\}}_{\Gamma}^{\circ}g(z)\;d z, $$ (6) where T is the positively oriented circle with center O and radius r diam of the complement of By (4), Theorem 14.15 can be applied to $F/a.$ It asserts that the diameter Since $(F/a)(U)$ is at most 4. Therefore diam $E\leq4|a|.$ $m\geq r.$ it follows that $$ |a|\geq_{4}^{r}. $$ (7)388 REAL AND coMPLEX ANALYSIS Since $\scriptstyle{\mathcal{G}}$ is a conformal mapping of Q onto $\mathrm{po};$ 1/|al),(7) shows that $$ |g(z)|<{\frac{4}{r}}\qquad(z\in\Omega) $$ (8) and since ${\Gamma}$ T is a path in $\Omega,$ of length 2mr, (6) gives $$ |b|<4r. $$ (9） If e D, then $|\zeta|<r,$ so (1),(8), and (9) imply $$ |Q|\leq{\frac{4}{r}}+5r\biggl({\frac{16}{r^{2}}}\biggr)<{\frac{100}{r}}. $$ (10) f This proves (Q2) , then $z g(z)=w F(w)/a;$ and since $w F(w)\to a$ as $w\to0,$ we have Fix<e D $z=F(w)$ zg(z)→1 as $z\to c o.$ Hence $\scriptstyle{\mathcal{G}}$ g has an expansion of the form $$ g(z)=\frac{1}{z-\zeta}+\frac{\lambda_{2}(\zeta)}{(z-\zeta)^{2}}+\frac{\lambda_{3}(\zeta)}{(z-\zeta)^{3}}+\cdot\cdot\cdot\;\;\;\;\;\;(|z-\zeta|>2r). $$ (11） Let ${\Gamma}_{\mathrm{0}}$ be a large circle with center at $0;(11)$ gives (by Cauchy's theorem) that $$ \lambda_{2}(\zeta)=\frac1{2\pi i}\left(\right)_{\Gamma_{0}}(z-\zeta)g(z)\;d z=b-\zeta. $$ (12） Substitute this value of 入,(G) into (11). Then (1) shows that the function $$ \varphi(z)=\left[Q(\zeta,z)-{\frac{1}{z-\zeta}}\right]\!(z-\zeta)^{3} $$ (13） is bounded as z→0. Hence $\varphi$ has a removable singularity at o.If $z\in\Omega\frown D.$ then $|z-\zeta|<2r,$ so (2) and (13) give $$ |\,\varphi(z)\,|\,<8r^{3}\,|\,Q(\zeta,\,z)\,|\,+\,4r^{2}\,<\,1,000r^{2}. $$ (14) By the maximum modulus theorem,(14) holds for all z e Q. This proves (3) /// 20.3 Lemma Suppose $f\in C_{c}^{\prime}(R^{2}),$ the space of all continuously differentiable functions in the plane, with compact support. Put $$ {\bar{\partial}}={\frac{1}{2}}\left({\frac{\partial}{\partial x}}+i{\frac{\partial}{\partial y}}\right). $$ (1） Then the following“Cauchy formula”holds: $$ f(z)=\,-\,{\frac{1}{\pi}}\left|\right|\!\!\int_{R^{2}}{\frac{(\bar{\partial}f)(\zeta)}{\zeta-z}}\,d\xi\,\,d\eta\qquad(\zeta=\xi+i\eta). $$ (2)UNIFORM APPROXIMATION BY POLYNOMIALS 389 PRoOF This may be deduced from Green's theorem. However, here is a simple direct proof: Put p(r, 0) = f(z + re"), r > 0,8 real. I $[\zeta=z+r e^{i\theta},$ the chain rule gives $$ (\bar{\partial}f)(\zeta)={\frac{1}{2}}\,e^{i\theta}\left[{\frac{\partial}{\bar{\partial}r}}+{\frac{i}{r}}\,{\frac{\partial}{\bar{\partial}\theta}}\right]\varphi(r,\theta). $$ (3） The right side of (2) is therefore equal to the limit, as $\scriptstyle\epsilon\to0.$ of $$ -\,\frac{1}{2\pi}\,\biggr)_{\epsilon}^{\;\infty}\,\displaystyle{\int_{0}^{2\pi}\left(\frac{\partial\varphi}{\partial r}+\frac{i}{r}\,\frac{\partial\varphi}{\partial\theta}\right)}\,d\theta\,d r. $$ (4） For each $\ r>0,$ $\varphi$ is periodic in (, with period 2z. The integral of Op/00 is therefore O, and (4) becomes $$ -{\frac{1}{2\pi}}\left[\right]_{0}^{2\pi}d\theta\,\left[\right]_{\epsilon}^{\left(\infty\right)}{\frac{\hat{\omega}\varphi}{\hat{\partial}r}}\,d r={\frac{1}{2\pi}}\int_{0}^{2\pi}\!\varphi(\epsilon,\,\theta)\,d\theta. $$ (5） Ase→0,p(e,0)→f(z) uniformly. This gives (2) // We shall establish Tietze's extension theorem in the same setting in which we proved Urysohn's lemma, since it is a fairly direct consequence of that lemma. 20.4 Tietze's Extension Theorem Suppose ${\cal K}\,\,$ is a compact subset of a locally compact Hausdorff space $X,$ and fe C(K). Then there exists an $F\in C_{c}(X)$ such that F(x) =f(x) for all $x\in K.$ (As in Lusin's theorem, we can also arrange it so that $\|F\|_{X}=\|f\|_{K}.$ PROOF Assume $\boldsymbol{\f}$ C is real, -1≤f≤1. Let $\mathbf{}W$ be an open set with compact closure so that $K\subset W.$ Put $$ K^{*}=\{x\in K:f(x)\geq{}\}\},\ \ \ \ \ K^{-}=\{x\in K:f(x)\leq-{}\}\} $$ (1） Then $K^{+}$ and $K^{-}$ are disjoint compact subsets of $\textstyle W.$ As a consequence of $\textstyle W.$ Urysohn's lemma there is a function $f_{1}\in C_{c}(X)$ such that $f_{1}(x)={\frac{1}{3}}$ on $K^{+}\!\!$ $f_{1}(x)=-{\frac{1}{3}}$ on $K^{-},\,-{\textstyle\frac{1}{3}}\leq f_{1}(x)\leq{\textstyle\frac{1}{3}}$ for al $x\in X_{\circ}$ and the support ${\mathfrak{o l}}/{\mathfrak{i}}$ lies in Thus $$ |f-f_{1}|\leq_{3}\mathrm{on}\ K,\qquad|f_{1}|\leq_{3}\mathrm{on}\ X. $$ (2) Repeat this construction with $f-f_{1}$ in place of $f\colon$ :： There exists an $\ f_{2}$ ∈ $C_{c}(X),$ with support in $W,$ so that $$ |f-f_{1}-f_{2}|\leq(\O_{3})^{2}{\mathrm{~on~}}K,\qquad|f_{2}|\leq\underline{{{3}}}\cdot{\frac{2}{3}}{\mathrm{~on~}}X. $$ (3) In this way we obtain functions f,∈ C,(X), with supports in W, such that $$ |f-f_{1}-\cdot\cdot\cdot-f_{n}|\leq(\O_{3}^{2})^{n}\ \mathrm{on}\ K,\qquad|f_{n}|\leq\textstyle\frac{1}{3}\cdot(\textstyle{\frac{2}{3}})^{n-1}\ \mathrm{ol}\quad $$ n X. (4)390 REAL AND coMPLEX ANALYSIS Put $F=f_{1}+f_{2}+f_{3}+\cdots.$ $X.$ Hence ${\mathbf{}}F$ is continuous. Also, the support of on $K,$ and it converges uniformly on By (4), the series converges to $\boldsymbol{\f}$ lies ${\mathbf{}}F$ in ${\overline{{W}}}.$ // Mergelyan's Theorem 20.5 Theorem If $\textstyle K$ is a compact set in the plane whose complement is con- which is holomorphic in the nected, if is a contimuous complex function on $\textstyle K$ interior of K,andi $\scriptstyle\epsilon\;>_{0},$ then there exists a polynomial ${\mathbf{}}P$ such tha |f(z)- P(z)|<e for all z e K. If the interior of $\textstyle K$ is empty, then part of the hypothesis is vacuously satis fied, and the conclusion holds for every fe C(K). Note that ${\cal K}\,\,$ need not be con nected. PROOF By Tietze's theorem, $\boldsymbol{\f}$ can be extended to a continuous function in the plane, with compact support. We fix one such extension, and denote it again byJ let $\scriptstyle{\omega(0)}$ be the supremum of the numbers For any $\delta>{\mathfrak{o}}.$ $$ |f(z_{2})-f(z_{1})| $$ where $z_{1}$ and $z_{2}$ are subject to the condition $|z_{2}-z_{1}|\leq\delta$ Since $\boldsymbol{\mathsf{f}}$ is uni- formly continuous, we have $$ \operatorname*{lim}_{\delta\lnot\Phi}\omega(\delta)=0. $$ (1） From now on, $\delta$ will be fixed. We shall prove that there is a polynomial ${\mathbf{}}P$ such that $$ |f(z)-P(z)|<10.000\omega(\delta)\qquad(z\in K). $$ (2) By (1), this proves the theorem for all z Our first objective is the construction of a function $\Phi\in C_{c}^{\prime}(R^{2}),$ such that l f(z)- 0(z)| ≤ 0(6), (3) $$ |(\hat{\sigma}\Phi)(z)|<\frac{2\omega(\delta)}{\delta}, $$ (4） and $$ \Phi(z)=\,-\,\frac{1}{\pi}\left[\right]\! [\frac{(\bar{d}\Phi)(\zeta)}{\zeta-z}\,d\zeta\,\,d\eta\qquad(\zeta=\zeta+i\eta), $$ (5)UNIFORM APPROxIMATION BY POLYNOMIALS 391 where $X$ is the set of all points in the support of p whose distance from the complement of $\textstyle K$ does not exceed 6.(Thus $X$ contains no point which is“ far within”K.) We construct $\Phi$ as the convolution of f with a smoothing function A. Put $a(r)=0$ if $\ r>\delta_{*}$ put $$ a(r)=\frac{3}{\pi\delta^{2}}\left(1-\frac{r^{2}}{\delta^{2}}\right)^{2}\,\qquad(0\leq r\leq\delta), $$ (6) and define $$ A(z)=a(|z|) $$ (7) for all complex z. It is clear that $A\in C_{c}^{\prime}(R^{2})$ We claim that $$ \begin{array}{c}{{\left.\displaystyle\int_{R^{2}}\right|A=1,}}\\ {{\left\{\displaystyle\int_{R^{2}}^{} \{\displaystyle\frac{1}{\delta A}=0,}}\\ {{ \{\displaystyle\frac{1}{\delta A}\right\}=\frac{24}{15\delta}<;}}\end{array} $$ 天 (10) (8) (9） The constants are so adjusted in (6) that (8) holds. (Compute the integral in polar coordinates.)(9) holds simply because A has compact support. To compute (10), express ${\bar{\partial}}A$ in polar coordinates, as in the proof of Lemma 20.3, and note that $\partial A/\partial\theta=0,\left|\partial A/\partial r\right|=-a^{\prime}(r).$ Now define $$ \Phi(z)=\prod_{R^{2}}^{\cdot}f(z-\zeta)A(\zeta)\;d\zeta\;d\eta=\prod_{R^{2}}^{\cdot}A(z-\zeta)f(\zeta)\;d\zeta\;d\eta. $$ (11 Since f and A have compact support, so does O. Since $$ \Phi(z)-f(z)=\prod_{R^{2}}^{r}\left[f(z-\zeta)-f(z)\right]A(\zeta)\;d\xi\;d\eta $$ (12) and $A(\zeta)=0$ if $|\zeta|>\delta,$ (3) follows from （8). The difference quotients of $\scriptstyle A$ converge boundedly to the corresponding partial derivatives of $A,$ since392 REAL AND coMPLEX ANALYSIs $A\in C_{c}^{\prime}(R^{2}).$ Hence the last expression in(11) may be differentiated under the integral sign, and we obtain （ $$ \begin{array}{c}{{\mathrm{j}\Phi(z)=\displaystyle\int_{a^{\prime}}^{\left[ (\bar{\partial}A\right)(z-\sqrt{\langle\bar{A}A\rangle(\bar{\partial}\rangle\,d\bar{\gamma}\,d\eta}}}}\\ {{\mathrm{~}=\displaystyle\int_{a^{\prime}}^{ [\int(z-\sqrt{\langle\bar{A}A\rangle(\bar{\partial})\,d\bar{\gamma}\,d\bar{\psi}}}}}\end{array} $$ ds dr. (13) The last equality depends on (9). Now (10) and（13) give(4). If we write (13) with $\Phi_{x}$ and $\Phi_{y}$ in place of ,we see that $\Phi_{\mathrm{\scriptscriptstyleI}}$ and $\mathbf{\nabla}(5)$ will follow if we can $\Phi$ has continuous partial derivatives. Hence Lemma 20.3 applies to show that ${\hat{\sigma}}\Phi=0$ in ${\cal G},$ where ${\boldsymbol{G}}$ is the set of all $\scriptstyle{\varepsilon\in K}$ whose distance from the complement of $ .K$ exceeds 6. We shall do this by showing that $$ \Phi(z)=f(z)\qquad(z\in G); $$ (14) note that ${\bar{\partial}}f=0$ in ${\cal G},$ since $\boldsymbol{\mathsf{f}}$ is holomorphic there. (We recall that $\tilde{\partial}$ is the Cauchy-Riemann operator defined in Sec. 11.1.) Now if $\scriptstyle{\varepsilon\in G}$ j, then $z-\zeta$ is in the interior of ${\cal K}\,\,\,\,\,\,\,\,$ for all $\textstyle\left\{{\begin{array}{l l}{\left\{\begin{array}{l l}{\end{array}}\right.}\end{array}}\right.$ with $|\zeta|<\delta.$ The mean value property for harmo- nic functions therefore gives, by the first equation in (11), $$ \begin{array}{c}{{\Phi(z)=\displaystyle\int_{0}^{\delta}a(r)r~d r\ \displaystyle\int_{0}^{2\pi}f(z-r e^{i\theta})~d\theta}}\\ {{\ }}\\ {{=2\pi f(z)\displaystyle\int_{0}^{\delta}a(r)r~d r=f(z)\displaystyle\int_{R^{2}}^{\delta}A=f(z)}}\end{array} $$ (15) for all $\varepsilon\in G$ We have now proved (3), (4), and (5). The definition of $\textstyle X$ X shows that $\textstyle X$ is compact and that $X$ can be covered $K.$ by finitely many open discs $S^{2}-K.$ It follows that each D contains a compact con- whose centers are not in Since $D_{1},\ldots,D_{n},$ of radius $2\delta,$ can be joined to oo by a $g^{2}-k$ is connected, the center of each $\underline{{D}}_{i}$ polygonal path in that nected set $E_{j},$ of diameter at least 26, so that $S^{2}-E_{j}$ is connected and so $K\cap E_{j}=\mathcal{D}.$ $r=2\delta.$ There exist functions We now apply Lemma_20.2,withUNIFORM APPROxIMATION BY PoLYNOMIALS 393 $g_{j}\in H(S^{2}-E_{j})$ and constants $b_{j}$ so that the inequalities $$ |Q_{i}\xi,z\}|<\frac{30}{\delta}, $$ (16) $$ \left|Q_{j}(\zeta,z)-\frac{1}{z-\zeta}\right|<\frac{4,000\delta^{2}}{|z-\zeta|^{3}} $$ (17） hold for z味 $E_{j}$ and $\zeta\in D_{j},$ if $$ Q_{j}(\zeta,z)=g_{j}(z)+(\zeta-b_{j})g_{j}^{2}(z). $$ (18) Let Q be the complement of $E_{1}\cup\cdot\cdot\sim\ \cup E_{n},$ Then $\Omega$ is an open set which contains K. 2 $:i\leq n.$ Define $$ {\begin{array}{l l}{X_{1}=X\cap D_{1}\quad{\mathrm{~and~}}\quad X_{j}=(X\cap D_{j})-(X_{1}\setminus\cdots\setminus X_{j-1}),}\\ {-1}&{}\end{array}}\, $$ for Put $$ R(\zeta,z)=Q_{j}(\zeta,z)\qquad(\zeta\in X_{j},z\in\Omega) $$ (19) and $$ F(z)=\frac{1}{\pi}\left|\!\right>^{\!\left(\widetilde{\theta}\Phi\right)\!(\zeta)R(\zeta,\,z)\ d\xi\ d\eta\qquad(z\in\Omega). $$ (20) Since $$ F(z)=\sum_{j=1}^{n}{\frac{1}{\pi}}\prod_{x_{j}}^{\left(\right)}(\tilde{\phi}\Phi)(\zeta)Q_{j}(\zeta,z)\;d\zeta\;d\eta, $$ (21） Hence (18) shows that ${\mathbf{}}F$ is a finite linear combination of the functions ${\mathfrak{g}}_{j}$ and $g_{j}^{2}.$ $F\in H(\Omega).$ By (20),(4), and (5) we have $$ |F(z)-\Phi(z)|<{\frac{2\omega(\delta)}{\pi\delta}}\prod_{\chi}\left|\right.\left|\frac{}{}R(\zeta,z)-{\frac{1}{z-\zeta}}\right|d\zeta\;d\eta\qquad(z\in\Omega). $$ (22 Q(C, Observe that the inequalities （16) and $\left(\bigsqcup{}^{\mu}{\mathcal{J}}\right)$ are valid with ${\boldsymbol{R}}$ in place of $Q_{j}$ if $\zeta\in X$ and z e Q. For if $\zeta\in X$ then $\zeta\in X_{j}$ for some ${\dot{j}},$ and then R(, z)= $z_{\mathrm{{\scriptsize~Z}}}$ for all z e Q2. if Now $\mathbb{H}x:\in\Omega,$ put $\zeta=z+\rho e^{i\theta},$ and estimate the integrand in (22) by (16) $\rho<4\delta,$ by (17) if $4\delta\leq\rho.$ The integral in (22) is then seen to be less than the sum of $$ 2\pi\mid_{\delta}^{*4\delta}{\left({\frac{50}{\delta}}+{\frac{1}{\rho}}\right)\rho~d\rho}=808\pi\delta $$ (23)394 REAL AND coMPLEX ANALYSIs and $$ 2\pi\, |_{4\delta}^{*\circ}\frac{4.000\delta^{2}}{\rho^{3}}\,\rho\,\,d\rho=2.000\pi\delta. $$ (24) Hence (22) yields $$ |F(z)-\Phi(z)|<6,000\omega(\delta)\qquad(z\in\Omega). $$ (25) Since F∈ H(Q) $\kappa\in\Omega,$ and $\scriptstyle{g^{2}-K}$ is connected, Runge's theorem shows that ${\mathbf{}}F$ can be uniformly approximated on ${\cal K}\,\,\,$ by polynomials. Hence $({\mathbf{3}})$ and (25) show that (2) can be satisfied This completes the proof // One unusual feature of this proof should be pointed out. We had to prove that the given function $\boldsymbol{\f}$ is in the closed subspace P(K) of C(K).(We use the by O. But terminology of Sec.20.1.) Our frst step consisted in approximating $\boldsymbol{\mathsf{f}}$ this step took us outside $\scriptstyle P(K),$ since p was so constructed that in general $\ \Phi$ will not be holomorphic in the whole interior of $K.$ Hence $\Phi$ is at some positive distance from $\scriptstyle P(k).$ However, (25) shows that this distance is less than a constant multiple of $\scriptstyle4(0)$ [In fact, having proved the theorem, we know that this distance is at most D(6), by (3) rather than 6,000 o(6).] The proof of (25) depends on the inequality (4) and on the fact that ${\hat{\sigma}}\Phi=0$ in ${\cal G}.$ Since holomorphic functions $\varphi$ are characterized by ${\bar{\partial}}\varphi=0,(4)$ may be regarded as saying that $\Phi$ is not too far from being holomorphic, and this interpretation is confirmed by (25) Exercises 1 Extend Mergelyan's theorem to the case in which $\scriptstyle{\vec{g}}\,-\,{\vec{kappa}}$ has finitely many components: Prove that then every fe C(K) which is holomorphic in the interior of ${\cal K}$ can be uniformly approximated on ${\mathbf{}}K$ by rational functions. discs in $\boldsymbol{\mathit{U}}$ whose union ${\boldsymbol{y}}$ is dense in $U_{\mathrm{,}}$ J, such that $\Sigma r_{n}<\infty$ Put $K={\bar{U}}-V.$ Let ${\Gamma}$ and in the plane, by be the 2 Show that the result of Exercise 1 does not extend to arbitrary compact sets $\scriptstyle{\mathcal{K}}$ verifying the details of the following example. For $n=1,2.3$ ,.…, let $D_{n}{}_{-}\,D(\alpha_{n};\,r_{n})$ be disjoint open paths $\gamma_{n}\,$ $$ \Gamma(t)=e^{i t},\quad\quad\gamma_{n}(t)=\emptyset_{n}+r_{n}e^{i t},\quad\quad0\leq t\leq2\pi, $$ and define $$ L(f)=\int_{\Gamma}f(z)\,d z-\,\sum_{n=1}^{\infty}\int_{\gamma_{n}}^{t}f(z)\,d z\qquad(f\in C(K)). $$ Prove that $\boldsymbol{\mathit{L}}$ is a bounded linear functional on $C(K),$ prove that $L(R)=0$ for every rational function R whose poles are outside $K,$ and prove that there exists $\mathbf{an}f\in C(K)$ for which $L(f)\neq0.$ among 3 Show that the function $\scriptstyle{\mathcal{G}}$ constructed in the proof of Lemma 20.2 has the smallest supremum norm can therefore be replaced by $\mathbf{a}|I\in H(\Omega)$ such that $z f(z)\to1$ as $z arrow\infty.$ (This motivates the proof of the lemma)） Show also that $b=c_{\mathrm{0}}$ in that proof and that the inequality ${\boldsymbol{E}}.$ $|b|<4r$ 1b|<r. In fact, b lies in the convex hull of the setAPPENDIX HAUSDORFF'S MAXIMALITY THEOREM We shall first prove a lemma which, when combined with the axiom of choice, leads to an almost instantaneous proof of Theorem 4.21. that If 沙 is a collection of sets and $\Phi\subset{\mathcal{F}}.$ , we call $\Phi$ a subchain of $\mathcal{F}$ provided and $\mathcal{F}$ $\mathbf{\Phi}\Phi$ is totally ordered by set inclusion. Explicitly, this means that if $\scriptstyle4\;\epsilon\oplus$ $B\in\Phi_{s}$ then either $\scriptstyle A\leq B$ or ${\mathfrak{g}}\in{\mathfrak{A}}$ The union of all members of $\Phi$ will simply be referred to as the union of $\Phi.$ Lemma Suppose . ${\widehat{\mathcal{M}}}$ is a nonempty collection of subsets of a set $X$ such that the associates to each $A\in{\mathcal{F}}$ union of every subchain of F belongs to 多.Suppose $\scriptstyle{\mathcal{G}}$ is a function which $g(A)-A$ a set g(A) ∈ F such that $A\in g(A)$ and consists of at most one element. Then there exists an A∈F for which $g(A)=A.$ PRO0F Fix $A_{0}$ e 多. Call a subcollection ${\mathcal{F}}^{\prime}$ of $\mathcal{F}$ a tower if ${\mathcal{F}}^{\prime}$ has the fol- lowing three properties: (a) $A_{0}\in{\mathcal{F}}^{\prime}.$ belongs to ${\mathcal{F}}^{\prime}.$ (b) The union of every subchain of ${\mathcal{F}}^{\prime}$ (c） If A∈ F', then also $g(A)\in{\mathcal{F}}^{\prime}.$ $A\in{\mathcal{F}}$ such that $4_{0}\subset A,$ The family of all towers is nonempty. For if ${\mathcal{F}}_{1}$ is a tower. Let ${\mathcal{F}}_{0}$ be the intersection of $\widetilde{\mathcal{P}^{\prime}}\[$ is the collection of all then all towers. Then ${\mathcal{F}}_{0}$ is a tower the verification is trivial),but no proper subcollection of ${\mathcal{P}}_{0}$ is a tower. Also, $A_{0}\subset A\ \ {\mathrm{if}}\ A\in{\mathcal{F}}_{0}$ The idea of the proof is to show that ${\mathcal{F}}_{0}$ is a subchain of 多 395